---
title: "Corpus Masterclass 0"
author: "Mark Alfano"
date: "27 April 2025"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set working directory
```{r}
setwd("~/Documents/Philosophy/Experimental/R")
```

## Windows Users

Make sure you've got RTools installed (Windows Users)

[https://cran.r-project.org/bin/windows/Rtools/rtools45/rtools.html]

Windows only check with: 
```{r}
find.package("pkgbuild")
```

## Mac users

If you're having trouble, [https://quanteda.io/articles/pkgdown/parallelisation.html#for-macos-users]

## Linux users

Magical invocation: `sudo apt install build-essential gcc libtbb-dev liblapack-dev libopenblas-dev libpoppler-cpp-dev libfreetype6-dev libpng-dev libtiff5-dev libjpeg-dev libharfbuzz-dev libfribidi-dev libcurl4-openssl-dev gfortran`

Ia Ia Cthulhu F'tagn... 


# Install packages
```{r}
install.packages(c("readtext", "topicmodels", "dplyr", "tidyr", "tidytext", "ggplot2", "network", "igraph", "RCurl", "stringr", "tm", "textstem", "qdapDictionaries", "lsa", "LSAfun", "wordcloud", "gridExtra",  "reshape2", "rmarkdown", "knitr"))


install.packages(c("quanteda", "quanteda.textplots", "quanteda.textmodels", "quanteda.textstats"))
```

# install developer tools and special packages

## PC users: search for and install Rtools; restart R Studio; then run the following chunk

```{r}
install.packages("devtools")
library(devtools)
devtools::install_github("quanteda/quanteda.corpora")
devtools::install_github("kbenoit/quanteda.dictionaries")

```
