---
title: "PDF to TF-IDF"
author: "Brian Ballsun-Stanton"
date: "2025-06-19"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE}
require(pdftools)
require(quanteda)
require(quanteda.textstats)
```

# Reuse Our Functions from Yesterday

```{r existing_functions}
# Same function from yesterday - no changes!
pdf_to_dataframe <- function(pdf_file, doc_id="dissertation") {
  pdf_text <- pdf_text(pdf_file) %>%
    paste(collapse = " ")
  
  word_count <- length(strsplit(pdf_text, "\\s+")[[1]])
  print(paste("Extracted", word_count, "words from", pdf_file))
  
  if (word_count < 100) {
    stop("Document has less than 100 words.")
  }
  
  df <- data.frame(
    doc_id = doc_id,
    text = pdf_text,
    stringsAsFactors = FALSE
  )
  
  return(df)
}

# Also from yesterday
prepare_df <- function(df, make_corpus=TRUE, remove_stopwords=TRUE) {
  if (make_corpus) {
    corp <- corpus(df, text_field = "text")
  } else {
    corp <- df 
  }
  
  toks <- tokens(corp, remove_punct = TRUE)
  
  if (remove_stopwords) {
    toks <- tokens_remove(toks, stopwords('en')) %>%
      tokens_tolower()
  }
  
  dfm_result <- dfm(toks)
  return(dfm_result)
}
```

# Build Corpus from Directory - Our First Loop!

```{r build_corpus_loop}
# Get list of PDF files
pdf_files <- list.files("pdfs/", pattern = "*.pdf", full.names = TRUE)

# Initialize empty list to store dataframes
all_dfs <- list()

# Loop through each PDF file
for (i in 1:length(pdf_files)) {
  pdf_file <- pdf_files[i]
  
  # Extract just the filename for doc_id
  doc_name <- tools::file_path_sans_ext(basename(pdf_file))
  
  # Try to process this PDF
  tryCatch({
    df <- pdf_to_dataframe(pdf_file, doc_id = doc_name)
    all_dfs[[i]] <- df
    print(paste("Success:", doc_name))
  }, error = function(e) {
    print(paste("Failed:", doc_name, "-", e$message))
  })
}

# Combine all successful PDFs into one dataframe
baseline_corpus_df <- do.call(rbind, all_dfs[!sapply(all_dfs, is.null)])
print(paste("Built corpus with", nrow(baseline_corpus_df), "documents"))
```

# Run TF-IDF with Our New Corpus

```{r tfidf_analysis}
# Load dissertation
diss_df <- pdf_to_dataframe("Tarnowski_phdthesis.pdf", doc_id = "dissertation")

# Create DFMs
diss_dfm <- prepare_df(diss_df)
baseline_dfm <- prepare_df(baseline_corpus_df)

# Same TF-IDF function from yesterday
save_tfidf_terms <- function(target_dfm, baseline_dfm, output_file = "terms.txt") {
  combined_dfm <- rbind(target_dfm, baseline_dfm)
  tfidf_result <- dfm_tfidf(combined_dfm)
  top_terms <- topfeatures(tfidf_result[1,], 50)
  writeLines(names(top_terms), output_file)
  return(top_terms)
}

# Get distinctive terms
tfidf_terms <- save_tfidf_terms(diss_dfm, baseline_dfm)
```

# Apply Custom Dictionary with KWIC - Our Second Loop!

```{r dictionary_kwic_loop}
# Read our curated dictionary from file
# We edited terms.txt down to just the terms we want
curated_dictionary <- readLines("curated_terms.dic")
print(paste("Dictionary contains", length(curated_dictionary), "terms"))

# Create tokens for KWIC
all_text <- c(diss_df$text, baseline_corpus_df$text)
all_ids <- c(diss_df$doc_id, baseline_corpus_df$doc_id)
corp <- corpus(all_text, docnames = all_ids)
toks <- tokens(corp)

# Create output directory
if (!dir.exists("kwic_output")) {
  dir.create("kwic_output")
}

# Loop through dictionary terms
for (term in curated_dictionary) {
  print(paste("Generating KWIC for:", term))
  
  # Create KWIC
  kwic_result <- kwic(toks, pattern = term, window = 10)
  
  # Save as simple CSV
  if (nrow(kwic_result) > 0) {
    output_file <- paste0("kwic_output/kwic_", term, ".csv")
    write.csv(as.data.frame(kwic_result), output_file, row.names = FALSE)
    print(paste("  Found", nrow(kwic_result), "instances"))
  } else {
    print(paste("  No instances found"))
  }
}
```

# Making a nested loop to look at co-occurance of dictionary terms within our corpus

Our plan is to iterate on our dictionary and then, within that, iterate on our dictionary, such that we can see how each term speaks when in relation to each other term.

```{r nested_loop}

window_size = 10

cooc_df <- data.frame(
  term1 = character(),
  term2 = character(),
  count = integer(),
  example = character(),
  stringsAsFactors = FALSE
)

for (term1 in curated_dictionary){
  kwic1 <- kwic(toks, pattern = term1, window = window_size)
  
  if (nrow(kwic1) > 0){
    for (term2 in curated_dictionary){
      if (term1 != term2){
        contexts <- paste(kwic1$pre, kwic1$keyword, kwic1$post)
        matches <- grepl(pattern=term2, x=contexts, ignore.case = TRUE)
        count <- sum(matches)
        
        if (count > 0){
          first_example <- contexts[which(matches)[1]]
          
          cooc_df <- rbind(cooc_df,
                           data.frame(
                             term1 = term1,
                             term2 = term2,
                             count = count,
                             example = first_example, #substr(first_example,1,100),
                             stringsAsFactors = FALSE
                             
                           ))
        }
      }
    }
  }
}

write.csv(cooc_df, "kwic_output/cooccurances.csv", row.names=FALSE)

```
