---
title: "Corpus Masterclass 2"
author: "Mark Alfano"
date: "27 April 2025"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set working directory
```{r}
setwd("~/Documents/Philosophy/Experimental/R")
```

# load required packages
```{r message=FALSE}
require(readtext)
require(quanteda)
require(quanteda.textplots)
require(quanteda.textstats)
require(quanteda.textmodels)
require(quanteda.dictionaries)
require(quanteda.corpora)
require(topicmodels)
require(dplyr)
require(tidyr)
require(tidytext)
require(ggplot2)
require(network)
require(igraph)
require(gridExtra)
library(lexicon)
library(reshape2)
library(stringi)
```

# LIWCalike analysis

## We can examine various predefined constructs represented in each SOTU; standard liwc dictionary
```{r}
liwcdic <- dictionary(file = "liwcdict.dic")
LIWCanalysis <- liwcalike(data_corpus_sotu, 
                          liwcdic)
head(LIWCanalysis)
```

### clean up LIWCalike output (could do more carefully, but why bother with Martin van Buren? He was basically Dutch)
```{r}
LIWCanalysis <- separate(LIWCanalysis, 
                         docname, 
                         c('president', 'year'))

LIWCanalysis$president <- replace(LIWCanalysis$president, 
                                  LIWCanalysis$president == "van", "vanBuren")

LIWCanalysis$year <- replace(LIWCanalysis$year, 
                             LIWCanalysis$year == "Buren", 1838)
```

### Further cleaning

```{r}
LIWCanalysis <- LIWCanalysis %>% mutate_at(c(2:92), 
                                           as.numeric)
```

### old way of doing it, column by column... it works but it's very clunky
```{r}
LIWCanalysis$year <- as.numeric(LIWCanalysis$year) 
LIWCanalysis$Sixltr <- as.numeric(LIWCanalysis$Sixltr)
LIWCanalysis$Posemo <- as.numeric(LIWCanalysis$posemo)
LIWCanalysis$Anx <- as.numeric(LIWCanalysis$anx)
LIWCanalysis$Anger <- as.numeric(LIWCanalysis$anger)
LIWCanalysis$Sad <- as.numeric(LIWCanalysis$sad)
```

## Now we can get the word counts from all SOTUs
```{r}
ggplot(data=LIWCanalysis, 
       aes(x = year, y = WC)) + 
      geom_line() + # is a line the best way to plot this data?
      geom_point() + # maybe a histogram?
      theme(axis.text.x = element_text(angle = 90))
```

## And various emotions
```{r}
ggplot(data=LIWCanalysis, 
       aes(x = year, y = posemo)) + 
      geom_line() + 
      geom_point() + 
      theme(axis.text.x = element_text(angle = 90)) # what happens when we change the angle?
```

```{r}
ggplot(data=LIWCanalysis, 
       aes(x = year, y = anx)) + 
      geom_line() + geom_point() + 
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis, 
       aes(x = year, y = anger)) + 
      geom_line() + 
      geom_point() + 
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis, 
       aes(x = year, y = sad)) + 
      geom_line() + 
      geom_point() + 
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis, 
       aes(x = year, y = pronoun)) + 
      geom_line() + 
      geom_point() + 
      theme(axis.text.x = element_text(angle = 90))
```


### Can we do all of them with a for loop?

```{r}

# Brian HALP!


```


## We can also examine the moral values represented in each SOTU, using the e-mac dictionary
```{r}
macdvirtue <- dictionary(file = "macdvirtue.dic")
LIWCanalysis_mac <- liwcalike(data_corpus_sotu, macdvirtue)
```

### clean up LIWCalike output
```{r}
LIWCanalysis_mac <- separate(LIWCanalysis_mac, docname, c('president', 'year'))
LIWCanalysis_mac$president <- replace(LIWCanalysis_mac$president, LIWCanalysis_mac$president == "van", "vanBuren")
LIWCanalysis_mac$year <- replace(LIWCanalysis_mac$year, LIWCanalysis_mac$year == "Buren", 1838)
```

### Further cleaning

```{r}
LIWCanalysis_mac <- LIWCanalysis_mac %>% mutate_at(c(2:26), as.numeric)
```

### old way, very clunky...
```{r}
LIWCanalysis_mac$year <- as.numeric(LIWCanalysis_mac$year)
LIWCanalysis_mac$Family <- as.numeric(LIWCanalysis_mac$family)
LIWCanalysis_mac$Group <- as.numeric(LIWCanalysis_mac$group)
LIWCanalysis_mac$Reciprocity <- as.numeric(LIWCanalysis_mac$reciprocity)
LIWCanalysis_mac$Heroism <- as.numeric(LIWCanalysis_mac$heroism)
LIWCanalysis_mac$Deference <- as.numeric(LIWCanalysis_mac$deference)
LIWCanalysis_mac$Fairness <- as.numeric(LIWCanalysis_mac$fairness)
LIWCanalysis_mac$Property <- as.numeric(LIWCanalysis_mac$property)
```


## MAC virtues in SOTU speeches
```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = family)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = group)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = reciprocity)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = heroism)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = deference)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = fairness)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(data=LIWCanalysis_mac, 
       aes(x = year, y = property)) + 
      geom_line() + 
      geom_point() +
      theme(axis.text.x = element_text(angle = 90))
```


## Transform the LIWCalike output into an adjacency matrix.   

```{r}
LIWCanalysis_mac_simple <- LIWCanalysis_mac[,8:14] # subset out only the columns representing the moral elements

M <- data.matrix(LIWCanalysis_mac_simple) # transform to matrix
T <- t(M) # create transpose of the matrix
LIWCanalysis_mac_adj <- (T %*% M) # multiply transpose by original matrix to create adjacency matrix
```

## transform adjacency matrix into a network object.

```{r}
mac_network <- graph_from_adjacency_matrix(LIWCanalysis_mac_adj, 
                                           mode = "undirected", # why?
                                           weighted = TRUE, # why?
                                           diag = FALSE) # why?
```

## weight nodes (vertices) and edges of the network
```{r}
str <- strength(mac_network)
V(mac_network)$size <- str*1.5 # why 1.5?
E(mac_network)$width <- E(mac_network)$weight

```

## plot the LIWCalike network using a couple different layouts (fruchterman-reingold, circle).
```{r}
l <- layout_with_fr(mac_network)
plot(mac_network, 
     layout=l, 
     vertex.color="pink", # what other colors might we use? Why?
     vertex.frame.color="pink", 
     vertex.label.color="black")
```

```{r}
l <- layout_in_circle(mac_network)
plot(mac_network, 
     layout=l, 
     vertex.color="pink", 
     vertex.frame.color="pink", 
     vertex.label.color="black")
```
