---
title: "20250616-CorpusMasterclass-4.Rmd"
author: "Brian Ballsun-Stanton"
date: "2025-06-16"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set working directory
```{r}
setwd("~/people/Mark/2025-CorpusLinguistics/")
```

# Install new packages for today
```{r}
# internetarchive: Access to Internet Archive's API
# Documentation: https://docs.ropensci.org/internetarchive/
devtools::install_github("ropensci/internetarchive", build_vignettes = TRUE)

# stringr: Modern string manipulation (part of tidyverse)
# Documentation: https://stringr.tidyverse.org/
# We may have used this before, but let's be sure
install.packages("stringr")
```

# Load required packages
```{r message=FALSE}
# Core text analysis packages we've used before
require(quanteda)
require(quanteda.textplots)
require(quanteda.textstats)
require(tidytext)
require(dplyr)
require(ggplot2)

# New package for downloading from Internet Archive
# Documentation: https://docs.ropensci.org/internetarchive/
require(internetarchive)

# For handling strings and cleaning
# Documentation: https://stringr.tidyverse.org/
require(stringr)
```

# Today's Goal: Building a Corpus from Scratch

Today we'll learn how to:
1. Download historical texts from the Internet Archive
2. Extract metadata and organize our files properly  
3. Clean OCR'd text (Optical Character Recognition often has errors)
4. Use **functions** to avoid repeating ourselves
5. Use **loops** to process multiple texts efficiently
6. Analyze our corpus using TF-IDF (Term Frequency-Inverse Document Frequency)
7. Visualize how distinctive terms appear across documents

## A Note on Pipes (%>%)

You've been using `%>%` throughout this course. This is called a "pipe" and it means "take the result of the left side and use it as the first argument on the right side." Think of it as "and then..."

```{r}
# Without pipes (harder to read):
result <- tolower("HELLO WORLD")
result <- str_trim(result)
print(result)

# With pipes (easier to read - flows like a sentence):
result <- "HELLO WORLD" %>%
  tolower() %>%
  str_trim()
print(result)
```

# Part 1: Manual Download - Understanding Each Step

Let's download our first text: Norbert Wiener's "Cybernetics" (1948). We'll do this step-by-step to understand what's happening.

## Step 1: Search for the item
```{r}
# The identifier for Wiener's Cybernetics
wiener_id <- "wiener-1938"

# Get metadata about this item from the Internet Archive
# This asks the Internet Archive: "Tell me about this item"
item_metadata <- ia_get_items(wiener_id)

# Let's see what we got
print(item_metadata)
```

## Step 2: Extract the metadata we need
```{r}
# The metadata comes in a complex structure. Let's extract what we need:
# Title, creator (author), and date

# First, let's look at what metadata fields are available
names(item_metadata$metadata)
```

```{r}
# Now extract the specific fields we want
title <- item_metadata$metadata$title
creator <- item_metadata$metadata$creator  
date <- item_metadata$metadata$date

# Print what we found
print(paste("Title:", title))
print(paste("Creator:", creator))
print(paste("Date:", date))
```

## Step 3: Download the text file
```{r}
# Internet Archive stores multiple file formats. We want the plain text (.txt) version
# First, let's see what files are available
files_available <- ia_files(wiener_id)
print(files_available)
```

```{r}
# Find the .txt file
txt_files <- files_available %>%
  filter(str_detect(name, "\\.txt$"))  # $ means "at the end of the string"

print(txt_files)

# Get the filename
txt_filename <- txt_files$name[1]  # Take the first .txt file
print(paste("Text file:", txt_filename))
```

## Step 4: Actually download the file
```{r}
# Download the text content
text_content <- ia_download_text(wiener_id, txt_filename)

# Let's see the first 500 characters
cat(substr(text_content, 1, 500))
```

## Step 5: Create a proper filename with metadata
```{r}
# Extract year from date (take first 4 characters)
year <- substr(date, 1, 4)

# Clean up creator name (remove special characters)
creator_clean <- str_replace_all(creator, "[^A-Za-z0-9 ]", "")  # Keep only letters, numbers, spaces
creator_clean <- str_replace_all(creator_clean, " ", "_")  # Replace spaces with underscores

# Clean up title  
title_clean <- str_replace_all(title, "[^A-Za-z0-9 ]", "")
title_clean <- str_replace_all(title_clean, " ", "_")

# Create filename: Year_Author_Title_(identifier).txt
# We keep the identifier to ensure uniqueness
filename <- paste0(year, "_", creator_clean, "_", title_clean, "_(", wiener_id, ").txt")
print(paste("Filename:", filename))
```

## Step 6: Save the file
```{r}
# Create a directory for our corpus if it doesn't exist
if (!dir.exists("corpus_day4")) {
  dir.create("corpus_day4")
}

# Save the text
writeLines(text_content, file.path("corpus_day4", filename))
print(paste("Saved:", filename))
```

# Part 2: Creating a Function - Don't Repeat Yourself!

Now we've downloaded one file, but we have 4 more to get. We COULD copy and paste all that code 4 times... but that would be:
- Tedious
- Error-prone  
- Hard to maintain
- Not scalable

Instead, let's create a **function** - a reusable piece of code.

## What is a function?

A function is like a recipe:
- It has a name
- It takes ingredients (inputs/arguments)
- It follows steps (the code inside)
- It produces a result (output)

```{r}
# Here's our download function
# We're wrapping all the steps we just did into one reusable package

download_ia_text <- function(identifier) {
  # This function takes one input: an Internet Archive identifier
  # It downloads the text and saves it with proper metadata
  
  print(paste("Processing:", identifier))
  
  # Step 1: Get metadata
  item_metadata <- ia_get_items(identifier)
  
  # Step 2: Extract metadata fields
  title <- item_metadata$metadata$title
  creator <- item_metadata$metadata$creator  
  date <- item_metadata$metadata$date
  
  # Step 3: Find the text file
  files_available <- ia_files(identifier)
  txt_files <- files_available %>%
    filter(str_detect(name, "\\.txt$"))
  
  # Check if we found a text file
  if (nrow(txt_files) == 0) {
    warning(paste("No .txt file found for", identifier))
    return(NULL)  # Exit the function early
  }
  
  txt_filename <- txt_files$name[1]
  
  # Step 4: Download the text
  text_content <- ia_download_text(identifier, txt_filename)
  
  # Step 5: Create proper filename
  year <- substr(date, 1, 4)
  creator_clean <- str_replace_all(creator, "[^A-Za-z0-9 ]", "")
  creator_clean <- str_replace_all(creator_clean, " ", "_")
  title_clean <- str_replace_all(title, "[^A-Za-z0-9 ]", "")
  title_clean <- str_replace_all(title_clean, " ", "_")
  
  filename <- paste0(year, "_", creator_clean, "_", title_clean, "_(", identifier, ").txt")
  
  # Step 6: Save the file
  writeLines(text_content, file.path("corpus_day4", filename))
  
  print(paste("Successfully saved:", filename))
  
  # Return information about what we downloaded
  # This creates a one-row data frame with our metadata
  return(data.frame(
    identifier = identifier,
    year = as.numeric(year),
    author = creator,
    title = title,
    filename = filename,
    stringsAsFactors = FALSE
  ))
}
```

## Test our function
```{r}
# Let's test with Shannon's information theory paper
shannon_info <- download_ia_text("in.ernet.dli.2015.503815")
print(shannon_info)
```

# Part 3: Using Loops - Process Multiple Items

Now we have a function, but we still need to call it multiple times. Enter the **loop**!

## What is a loop?

A loop repeats code for each item in a list. Instead of:
```
download_ia_text("id1")
download_ia_text("id2")  
download_ia_text("id3")
```

We can write:
```
for (id in list_of_ids) {
  download_ia_text(id)
}
```

## Download our remaining texts
```{r}
# List of remaining identifiers
identifiers <- c(
  "MIND--COMPUTING-MACHINERY-AND-INTELLIGENCE",  # Turing
  "anarchy_Cyborg_Manifesto_Harroway",          # Haraway
  "elizabeth-anscombe-metaphysics-and-philosophy-of-mind"  # Anscombe
)

# Create an empty list to store our results
# Lists in R can hold different types of data
corpus_metadata <- list()

# We already have Wiener and Shannon, so let's add them
corpus_metadata[[1]] <- data.frame(
  identifier = "wiener-1938",
  year = 1948,
  author = "Norbert Wiener",
  title = "Cybernetics",
  filename = list.files("corpus_day4", pattern = "wiener", full.names = FALSE)[1],
  stringsAsFactors = FALSE
)

corpus_metadata[[2]] <- shannon_info

# Now loop through the remaining texts
# The loop says: "For each identifier in our list, do the following:"
for (i in 1:length(identifiers)) {
  # Download the text and get its metadata
  result <- download_ia_text(identifiers[i])
  
  # Add to our metadata list
  # i+2 because we already have 2 items
  corpus_metadata[[i + 2]] <- result
  
  # Pause for 2 seconds to be polite to Internet Archive's servers
  Sys.sleep(2)
}

# Combine all metadata into one data frame
# do.call applies a function (rbind) to a list of arguments
corpus_df <- do.call(rbind, corpus_metadata)
print(corpus_df)
```

# Part 4: Text Cleaning Function

OCR'd texts often have artifacts. Let's create a function to clean them.

## First, let's examine what we're dealing with
```{r}
# Read one of our downloaded files
sample_file <- file.path("corpus_day4", corpus_df$filename[1])
sample_text <- readLines(sample_file)

# Look at the first 20 lines
cat(sample_text[1:20], sep = "\n")
```

## Create a cleaning function
```{r}
clean_text <- function(text_lines) {
  # This function takes a vector of text lines and cleans them
  # Input: character vector (one element per line)
  # Output: single cleaned string
  
  # Step 1: Combine all lines into one string
  text <- paste(text_lines, collapse = " ")
  
  # Step 2: Basic cleaning
  text <- text %>%
    str_replace_all("\\s+", " ") %>%  # Replace multiple spaces with single space
    str_replace_all("- ", "") %>%      # Remove hyphenation at line breaks
    str_trim()                         # Remove leading/trailing whitespace
  
  return(text)
}

# Test our cleaning function
sample_cleaned <- clean_text(sample_text)
cat(substr(sample_cleaned, 1, 500))
```

# Part 5: Building Our Corpus

Now let's read all our texts and create a proper corpus object.

```{r}
# Initialize an empty list for our texts
texts <- list()

# Loop through each file and read it
for (i in 1:nrow(corpus_df)) {
  # Construct the full file path
  filepath <- file.path("corpus_day4", corpus_df$filename[i])
  
  # Read the file
  text_lines <- readLines(filepath)
  
  # Clean the text
  cleaned_text <- clean_text(text_lines)
  
  # Store in our list
  texts[[i]] <- cleaned_text
  
  # Give it a name based on the metadata
  names(texts)[i] <- paste(corpus_df$year[i], corpus_df$author[i])
}

# Create a quanteda corpus
# We're building this properly from the start so it can scale!
corpus_texts <- corpus(texts)

# Add our metadata to the corpus
docvars(corpus_texts) <- corpus_df

# Check what we have
summary(corpus_texts)
```

# Part 6: Tokenization and Identifying OCR Artifacts

Let's tokenize and see what OCR artifacts we need to clean.

```{r}
# Tokenize
toks <- tokens(corpus_texts, 
               remove_punct = TRUE,
               remove_numbers = TRUE)

# Create document-feature matrix
dfm_texts <- dfm(toks)

# Look at the most frequent features
topfeatures(dfm_texts, 50)
```

## Identify and remove OCR artifacts
```{r}
# Common OCR artifacts are:
# - Single letters (except "a" and "I")
# - Very long "words" (likely run-together text)
# - Repeated characters

# Let's see what single/double character tokens we have
short_tokens <- topfeatures(dfm_texts, 200)
short_tokens <- names(short_tokens[nchar(names(short_tokens)) <= 2])
print(short_tokens)

# Create custom stopwords combining English stopwords and OCR artifacts
custom_stop <- c(
  stopwords("en"),
  short_tokens[!short_tokens %in% c("a", "i")],  # Keep "a" and "I"
  "pg", "pp", "vol", "ed", "eds", "cf",          # Common academic abbreviations
  "ii", "iii", "iv", "v", "vi", "vii", "viii"   # Roman numerals
)

# Rebuild our dfm with custom stopwords
dfm_clean <- dfm_texts %>%
  dfm_remove(pattern = custom_stop) %>%
  dfm_trim(min_termfreq = 2)  # Remove terms that appear only once

# Check our cleaned top features
topfeatures(dfm_clean, 30)
```

# Part 7: TF-IDF Analysis

TF-IDF helps us find words that are distinctive to each document.

```{r}
# Calculate TF-IDF
dfm_tfidf <- dfm_tfidf(dfm_clean)

# For each document, find the top 10 TF-IDF terms
# This tells us what makes each text unique
for (i in 1:ndoc(dfm_tfidf)) {
  doc_name <- docnames(dfm_tfidf)[i]
  
  # Get TF-IDF scores for this document
  scores <- dfm_tfidf[i, ]
  
  # Sort and get top 10
  top_terms <- sort(scores, decreasing = TRUE)[1:10]
  
  cat("\n", doc_name, "- Distinctive terms:\n")
  print(top_terms)
}
```

## Extract top 2 terms per document for visualization
```{r}
# Create a data frame to store our top terms
top_terms_df <- data.frame()

for (i in 1:ndoc(dfm_tfidf)) {
  # Get document info
  doc_name <- docnames(dfm_tfidf)[i]
  doc_year <- docvars(dfm_tfidf)$year[i]
  doc_author <- docvars(dfm_tfidf)$author[i]
  
  # Get TF-IDF scores
  scores <- dfm_tfidf[i, ]
  
  # Get top 2 terms
  top_2 <- sort(scores, decreasing = TRUE)[1:2]
  
  # Add to our data frame
  for (j in 1:2) {
    top_terms_df <- rbind(top_terms_df, data.frame(
      term = names(top_2)[j],
      tfidf = as.numeric(top_2[j]),
      document = doc_name,
      year = doc_year,
      author = doc_author,
      rank = j,
      stringsAsFactors = FALSE
    ))
  }
}

print(top_terms_df)
```

# Part 8: Visualizing Term Distribution

Now let's see how each document's distinctive terms appear across ALL documents.

```{r}
# Get unique terms from our top terms
terms_to_plot <- unique(top_terms_df$term)
print(paste("Tracking these terms:", paste(terms_to_plot, collapse = ", ")))

# Create a color mapping for terms by their source document
term_colors <- top_terms_df %>%
  group_by(term) %>%
  slice(1) %>%  # Take first occurrence of each term
  select(term, document)

# Create keyword-in-context for all our terms
kwic_results <- list()

for (term in terms_to_plot) {
  kwic_results[[term]] <- kwic(toks, pattern = term)
}

# Combine all kwic results
all_kwic <- do.call(rbind, kwic_results)
```

## Create the lexical dispersion plot
```{r fig.width=10, fig.height=8}
# Sort corpus by year for chronological display
corpus_sorted <- corpus_texts[order(docvars(corpus_texts)$year)]

# Create tokens from sorted corpus
toks_sorted <- tokens(corpus_sorted, 
                      remove_punct = TRUE,
                      remove_numbers = TRUE)

# Plot lexical dispersion
# Each row is a term, each column is position in text
# Color intensity shows frequency
textplot_xray(
  kwic(toks_sorted, pattern = terms_to_plot),
  scale = "relative",  # Use relative position so different length texts are comparable
  sort = FALSE  # Keep chronological order
) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold")
  ) +
  labs(
    title = "Distinctive Terms Across Early Computing & Philosophy Texts",
    subtitle = "Terms sized by TF-IDF importance in their source documents",
    x = "Relative Position in Document",
    y = "Term"
  )
```

## Alternative visualization with document labels
```{r fig.width=12, fig.height=8}
# Create a more informative plot with document boundaries
# First, we need to track which terms come from which documents

# Create a color palette
library(RColorBrewer)
doc_colors <- brewer.pal(n = ndoc(corpus_texts), name = "Set3")

# Create a mapping of terms to their source document colors
term_source <- top_terms_df %>%
  group_by(term) %>%
  slice(1) %>%
  mutate(color = doc_colors[as.numeric(as.factor(document))])

# Create individual kwic for each term to maintain source info
plot_data <- data.frame()

for (i in 1:nrow(top_terms_df)) {
  term <- top_terms_df$term[i]
  source_doc <- top_terms_df$document[i]
  
  # Get kwic for this term
  kwic_term <- kwic(toks_sorted, pattern = term)
  
  if (nrow(kwic_term) > 0) {
    # Add source document info
    kwic_df <- as.data.frame(kwic_term)
    kwic_df$source_document <- source_doc
    kwic_df$term <- term
    
    plot_data <- rbind(plot_data, kwic_df)
  }
}

# Create a custom plot
library(ggplot2)
ggplot(plot_data, aes(x = from, y = term, color = source_document)) +
  geom_point(alpha = 0.7, size = 2) +
  facet_wrap(~ docname, scales = "free_x", ncol = 1) +
  scale_color_brewer(palette = "Set3", name = "Term Source") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    title = "Cross-Pollination of Distinctive Terms in Early Computing Texts",
    subtitle = "Each term is colored by the document where it has highest TF-IDF score",
    x = "Position in Document",
    y = "Term"
  )
```

# Summary: What We've Learned

Today we've built a complete corpus analysis pipeline using:

1. **Functions** - Reusable code blocks that:
   - Take inputs (arguments)
   - Process data
   - Return outputs
   - Save us from repetition

2. **Loops** - Code that repeats for each item in a collection:
   - `for` loops iterate through lists
   - Process multiple files efficiently
   - Build scalable workflows

3. **Good Design Principles**:
   - Store data in scalable structures (lists, data frames)
   - Build metadata management into our workflow
   - Create clean, reusable functions
   - Document our code with comments

4. **Practical Skills**:
   - Downloading texts from Internet Archive
   - Cleaning OCR artifacts
   - TF-IDF analysis for finding distinctive terms
   - Visualizing term distribution across documents

## What's Next?

In the afternoon session, we'll extend this pipeline to:
- Download PDFs from PhilPapers
- Convert PDFs to text
- Add them to our existing corpus
- Re-run our analysis without changing our code!

The beauty of building it right from the start is that our code is ready to scale.

## Save Our Work
```{r}
# Save our corpus metadata for the afternoon session
write.csv(corpus_df, "corpus_day4_metadata.csv", row.names = FALSE)

# Save the cleaned corpus object
saveRDS(corpus_texts, "corpus_day4_texts.rds")

print("Morning session complete! Your corpus is ready for expansion.")
```
