---
title: "Building a Dictionary"
author: 
- "Brian Ballsun-Stanton"
date: "2025-06-19"
output: html_document
editor_options: 
  chunk_output_type: console
---
# AI Use Disclosure:

- [https://claude.ai/share/13e8baa5-4460-44b4-9411-fa3839615280]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Set working directory
```{r}
setwd("~/people/Mark/2025-CorpusLinguistics/")
```

# Load required packages
```{r requires, message=FALSE}
require(pdftools)      # for PDF text extraction
require(quanteda)      # for corpus analysis
require(quanteda.corpora)  # for baseline corpora
require(quanteda.textstats)  # for TF-IDF
require(hunspell)      # for spell checking (optional)
```

# Introduction: Functions and If Statements

## What are functions?

Yesterday, we used a loop to automatically create multiple plots. Today, we'll learn about **functions** - reusable chunks of code that:
- Take inputs (called "arguments" or "parameters")
- Do something with those inputs
- Return an output

Think of a function like a recipe:
- **Ingredients** (inputs): What you put in
- **Instructions** (the function body): What you do with the ingredients
- **Final dish** (output): What you get back

## What are if statements?

If statements let your code make decisions. They check if something is true, and then do different things based on the answer. Like:
- IF it's raining, THEN take an umbrella
- IF the PDF is empty, THEN stop and warn the user

# Function 1: Converting PDF to a Data Frame

Let's build our first function to extract text from a PDF and put it in a format that quanteda can understand.

```{r pdf_df_function}
# Function definition starts with the function name and what goes in the parentheses
pdf_to_dataframe <- function(pdf_file, doc_id = "dissertation") {
  # pdf_file: the path to your PDF file
  # doc_id: what we'll call this document (default is "dissertation")
  
  # Step 1: Extract text from PDF
  # pdf_text() returns a vector with one element per page
  text <- pdf_text(pdf_file)
  
  # Step 2: Combine all pages into one long text
  # collapse = " " means put a space between each page
  text <- paste(text, collapse = " ")
  
  word_count <- length(strsplit(text, "\\s+")[[1]])
  print(paste("Extracted", word_count, "words from", pdf_file))

  # Second IF statement: Check if we have a reasonable number of words
  if (word_count < 100) {
    stop("PDF extraction suspicious - document has less than 100 words!")
  }
  
  
  # Step 3: Create a data frame that matches quanteda's expected format
  # quanteda corpora always have at least doc_id and text columns
  df <- data.frame(
    doc_id = doc_id,           # document identifier
    text = text,               # the actual text content
    stringsAsFactors = FALSE   # keep text as text, not categories
  )
  
  # Step 4: Send the data frame back to whoever called the function
  return(df)
}

# Test the function

test <- pdf_to_dataframe("20250603-BBS-JT-TeachingTheUnknown.pdf", doc_id="TeachingUnkown")

```

**Key concepts:**
- `function()` creates a new function
- Everything inside `{ }` is what the function does
- `return()` sends the result back
- `doc_id = "dissertation"` sets a default value - if you don't specify doc_id, it uses "dissertation"



# Function 2: Creating a Document-Feature Matrix with Conditional Logic

Now let's create a function that prepares our text for analysis, with an option to remove stopwords or not.

```{r}
prepare_dfm <- function(df, make_corpus=TRUE, remove_stopwords = TRUE) {
  # df: a data frame with doc_id and text columns
  # remove_stopwords: should we remove common words like "the", "and"? (default: yes)
  
  # Step 1: Convert data frame to corpus object
  # text_field tells corpus() which column contains the text
  if (make_corpus){
    corp <- corpus(df, text_field = "text")
  } else {
    corp <- df
  }
  
  # Step 2: Tokenize (split into words) and remove punctuation
  # tokens() breaks text into individual words
  # remove_punct = TRUE strips out . , ! ? etc.
  toks <- tokens(corp, remove_punct = TRUE)
  
  # Step 3: HERE'S OUR IF STATEMENT!
  # This is where the function makes a decision
  if (remove_stopwords) {
    # If remove_stopwords is TRUE, remove common English words
    # This helps focus on meaningful content words
    toks <- tokens_remove(toks, stopwords("en"))
    
    # We could add more cleaning here if needed:
    # toks <- tokens_tolower(toks)  # make everything lowercase
  }
  # If remove_stopwords is FALSE, we skip this step entirely
  
  # Step 4: Convert tokens to document-feature matrix
  # This creates a table: rows = documents, columns = words, cells = counts
  dfm_result <- dfm(toks)
  
  # Step 5: Return the DFM
  return(dfm_result)
}

test_dfm <- prepare_dfm(test)
```

**The IF statement explained:**
- `if (condition) { do something }` 
- If the condition is TRUE, run the code inside `{ }`
- If the condition is FALSE, skip that code
- This lets us toggle stopword removal on/off by changing one parameter

# Function 3: TF-IDF Analysis and Saving Results

TF-IDF helps us find words that are distinctive to our document compared to a baseline corpus.

```{r}
save_tfidf_terms <- function(target_dfm, baseline_dfm, output_file = "distinctive_terms.txt") {
  # target_dfm: the DFM of our document (e.g., dissertation)
  # baseline_dfm: the DFM of comparison documents (e.g., Guardian articles)
  # output_file: where to save the results
  
  # Step 1: Combine our document with the baseline corpus
  # rbind() stacks DFMs vertically (row bind)
  # Now row 1 = our document, rows 2+ = baseline documents
  combined_dfm <- rbind(target_dfm, baseline_dfm)
  
  # Step 2: Calculate TF-IDF scores
  # TF-IDF = Term Frequency × Inverse Document Frequency
  # High score = word is frequent in our doc but rare in others
  tfidf_result <- dfm_tfidf(combined_dfm)
  
  # Step 3: Extract top 50 terms from OUR document (row 1)
  # [1,] means "first row, all columns"
  # topfeatures() returns words sorted by their TF-IDF scores
  top_terms <- topfeatures(tfidf_result[1,], 50)
  
  # Step 4: Save just the word list to a text file
  # names(top_terms) gets the words (not the scores)
  # One word per line makes it easy to edit
  writeLines(names(top_terms), output_file)
  
  # Step 5: Also return the terms so we can see them
  return(top_terms)
}

dfm_sotu <- prepare_dfm(data_corpus_sotu, make_corpus = FALSE)
test_terms <- save_tfidf_terms(test_dfm, dfm_sotu)
```

# Key Takeaways

1. **Functions** let us package code for reuse:
   - Define once, use many times
   - Make complex tasks simple
   - Share code with others

2. **If statements** let code make decisions:
   - Check conditions
   - Do different things based on the result
   - Make functions flexible

3. **TF-IDF** helps find distinctive vocabulary:
   - Compares your document to a baseline
   - High scores = distinctive to your document
   - Still requires human judgment to select meaningful terms

4. **Dictionary building** combines computation and expertise:
   - Start with data-driven suggestions (TF-IDF)
   - Apply domain knowledge to select terms
   - Create reusable resources for analysis

# Next Steps

Tomorrow we'll expand on this foundation:
- Download multiple documents in a loop
- Process mixed file types (PDF and TXT)
- Build larger corpora from online sources
- Create more sophisticated visualizations

Remember: Functions and if statements are the building blocks that let us scale from analyzing one document to analyzing hundreds!