---
title: "SOTU Wordclouds"
author: "Brian Ballsun-Stanton"
date: "2025-06-17"
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 20250617 1501

* Started adding requirements.

```{r requirements, include=FALSE}
require(quanteda)
require(quanteda.corpora)
require(quanteda.textplots)
require(ggplot2)
require(stopwords)
```

## 20250617 1514

* Teaching pipes
* Making a DFM out of the SOTU (in variable `dfmat_sotu`)

```{r dfm_sotu, message=FALSE, echo=FALSE}
# Document feature matrix of quanteda corpus, State of the union

dfmat_sotu <- tokens(data_corpus_sotu,
                    remove_punct = TRUE) %>%
              dfm() %>%
              dfm_remove(pattern = stopwords("en")) %>%
              dfm_trim(min_docfreq = 3,
                       min_termfreq = 5)
  

```

## Make. Wordcloud?

* Let's make the wordcloud now, because we're presenting to a manager.

```{r wordcloud, echo=FALSE}
set.seed(1)

textplot_wordcloud(dfmat_sotu,
                   min_count = 600,
                   random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2")
                   ) # 8 is n, the set of different colours.
```


## Custom colours in wordcloud

* having searched, I found the 12 bit palette https://iamkate.com/data/12-bit-rainbow/ and 
* https://r-graph-gallery.com/ggplot2-color.html

ggplot2 natively supports several methods to customize the color palette:

p <- ggplot(iris,
  aes(x=Sepal.Length, y=Sepal.Width, color=Species)) +
  geom_point(size=6)
        
p + scale_color_manual(values=c("#69b3a2", "purple", "black"))


Hypothesis: I can set color = scale_color_manual(values blah for the wordcloud)

```{r wordcloud_ex1, echo=FALSE}
# set.seed(1)
# 
# textplot_wordcloud(dfmat_sotu,
#                    min_count = 600,
#                    random_order = FALSE,
#                    rotation = .25,
#                    color = scale_color_manual(values=c("#69b3a2", "purple", "black")) #RColorBrewer::brewer.pal(8, "Dark2")
# Error in color[cc] : object of type 'environment' is not subsettable
                   # ) # 8 is n, the set of different colours.
```

Experiment didn't work.

## 20250618 0944

Experiment 2, try a standard vector.

```{r wordcloud_ex2, echo=FALSE}
set.seed(1)

textplot_wordcloud(dfmat_sotu,
                   min_count = 600,
                   random_order = FALSE,
                   rotation = .25,
                   color = c("#69b3a2", "purple", "black")
)
```

Experiment worked! Now to get the 12-bit colours!


```{r wordcloud_rainbow, echo=FALSE}
#| fig-cap: "A good rainbow SOTU render"
#| fig-alt: "High frequency words from the SOTU"

textplot_wordcloud(dfmat_sotu,
                   min_count = 600,
                   random_order = FALSE,
                   rotation = .25,
                   color = c("#817", "#a35", "#c66", "#e94", "#ed0", "#9d5", "#4d8", "#2cb", "#0bc", "#09c", "#36b", "#639")
)
```

Bug: "Duplicate label wordcloud" -- when making chunks, make sure to change the labels

According to [@smithEducatingRiskHow2020] rainbows are neat...